\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,graphicx,
enumerate,fancyhdr}
\usepackage[letterpaper, margin=0.9in]{geometry}
\usepackage[scr=boondoxo]{mathalpha}

% ordered lists have letters instead of numbers
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\labelenumi{(\theenumi)}

% use hyperref for links but don't draw the ugly boxes around links
\usepackage[hidelinks]{hyperref}

% command to create underlined link to outside websites
\newcommand{\hrefunderline}[2]{\underline{\href{#1}{#2}}}

% font to use for collections
\newcommand{\col}[1]{\mathscr{#1}}
% font to use for random variables
\newcommand{\rv}[1]{\mathsf{#1}}
% the probability metric
\newcommand{\p}{\mathbb{P}}
% expectation value
\newcommand{\ex}{\mathbb{E}\,}
% indicator functions
\newcommand{\ind}[1]{\mathbf{1}_{#1}}
% the Borell sigma algebra on the reals
\newcommand{\bor}{\col{B}}
% power sets
\newcommand{\pset}[1]{P\left(#1\right)}
% command to emphasize the name of something defined
\newcommand{\defname}[1]{\underline{#1}}
% almost sure convergence
\newcommand{\asto}{\xrightarrow{\text{a.s.}}}
% convergence in probability
\newcommand{\pto}{\xrightarrow{\p}}
% convergence in distribution
\newcommand{\disto}{\xrightarrow{\text{dist}}}
% convergence in expectation
\newcommand{\exto}{\xrightarrow{L^1}}


% symbols for common sets
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\CC}{\mathbb{C}}

% for congruences
\newcommand{\mmod}[1]{\;(\operatorname{mod} {#1})}

% new theorem style with boldface title and italic content
\newtheoremstyle{step}%            % Name
  {}%                                     % Space above
  {}%                                     % Space below
  {\itshape}%                           % Body font
  {}%                                     % Indent amount
  {\itshape}%                          % Theorem head font
  {:}%                                    % Punctuation after theorem head
  { }%                                    % Space after theorem head, ' ', or \newline
  {}%                                     % Theorem head spec (can be left empty, meaning `normal')

% new theorem style for a gap in the notes
\newtheoremstyle{gap}%            % Name
  {}%                                     % Space above
  {}%                                     % Space below
  {\itshape}%                           % Body font
  {}%                                     % Indent amount
  {\itshape}%                          % Theorem head font
  {!!!}%                                    % Punctuation after theorem head
  {    }%                                    % Space after theorem head, ' ', or \newline
  {}%                                     % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}

\theoremstyle{remark}
\newtheorem{example}[theorem]{Example}
\newtheorem*{remark}{Remark}

\theoremstyle{step}
\newtheorem{step}{Step}[subsection]
\renewcommand{\thestep}{\arabic{step}}

\theoremstyle{gap}
\newtheorem*{gap}{Gap}

\begin{document}

\title{Probability Notes}
\author{Duncan MacIntyre}
\date{\today}
\maketitle
\tableofcontents
\bigskip
\newpage

\section{Introduction to probability theory}

\subsection{Some measure theory}

\begin{gap}
The discussion of topologies should probably be removed.
\end{gap}


\begin{definition}
Let \(\Omega\) be a set. A \defname{topology} on \(\Omega\) is a collection \(\col{A} \subset \pset{\Omega}\) that is closed under unions and finite intersections with \(\Omega, \emptyset \in \col{A}\). For \(E \subset \Omega\), if \(E \in \col{A}\), we call \(E\) open, and if \(E^C \in \col{A}\), we call \(E\) closed.
\end{definition}

For example, if \((X, d)\) is a metric space, \(\col{A} = \left\{E | E = \cup_i N_{r_i} \left(x_i\right)\right\}\) is a topology on \(X\).

\begin{gap} The example should be proven.\end{gap}

\begin{definition}
Let \(\Omega\) be a set and \(\col{S} \subset \pset{\Omega}\). Then the \defname{topology generated by \(\col{S}\)} is
\[\col{A} = \left\{E \subset \Omega : E \text{ is a union of finite intersections of sets in } \col{S}.\right\}\]
\end{definition}

\begin{gap} The definition should be more explicit.\end{gap}

\begin{gap} Examples and Furstenberg's theorem should be added here \end{gap}

\begin{definition}
Let \(\Omega\) be a set. An \defname{algebra} is a collection \(\col{A} \subset \pset{\Omega}\) that is closed under finite unions and compliments with \(\Omega \in \col{A}\). If an algebra is also closed under countable unions, we call it a \defname{\(\sigma\)-algebra}.
\end{definition}

\begin{theorem}
Algebras are closed under finite intersections and \(\sigma\)-algebras are closed under countable intersections.
\end{theorem}

\begin{proof}
Let \(\col{A}\) be an algebra and let \(A_1, \ldots, A_n \in \col{A}\). Then \(A_1^C, \ldots, A_n^C \in \col{A}\), so \(\left(\cap_{i=1}^n A_i\right)^C = \cup_{i=1}^n A_i^C \in \col{A}\), so \(\cap_{i=1}^n A_i \in \col{A}\). The proof for \(\sigma\)-algebras is similar.
\end{proof}

\begin{definition}
Let \(\col{f}\) be a \(\sigma\)-algebra on a set \(\Omega\). A \defname{measure} on \(\col{f}\) is a function \(\mu:\col{f} \to [0, \infty]\) such that \(\mu(\emptyset) = 0\) and for all disjoint \(\left(A_i\right)_{i\in \NN} \in \col{f}\) we have \(\mu\left(\cup_i A_i\right) = \sum_i \mu\left(A_i\right)\). We call \((\Omega, \col{f}, \mu)\) a \defname{measure space}.
\end{definition}

\begin{gap}
Add examples of measures and probability measures.
\end{gap}

\begin{theorem}\label{thm.intersection-algebras}
An intersection of \(\sigma\)-algebras is also a \(\sigma\)-algebra.
\end{theorem}

\begin{proof}
Let \((\col{A}_i)\) be uncountably many \(\sigma\)-algebras. Let \(A \in \cap \col{A}_i\). Then \(A^C \in \col{A}_i\) for all \(\col{A}_i\), so \(A^C \in \cap \col{A}_i\). Similarly, let \((A_j)_{j \in \NN} \in \cap \col{A}_i\). Then \(\cup_j A_j \in \col{A_i}\) for all \(\col{A}_i\), so \(\cup_j A_j \in \cap \col{A_i}\).
\end{proof}

Theorem \ref{thm.intersection-algebras} motivates the following definition.

\begin{definition}
Let \(\col{A} \subset P\left(\Omega\right)\). The \defname{\(\sigma\)-algebra generated by \(\col{A}\)} is the intersection of all \(\sigma\)-algebras that contain \(\col{A}\). We denote it by \(\sigma(\col{A})\).
\end{definition}

We can think of \(\sigma(\col{A})\) as the smallest \(\sigma\)-algebra containing \(\col{A}\).

\begin{notation}
A Borel \(\sigma\)-algebra is a \(\sigma\)-algebra generated by a topology. We will use \(\bor\) to denote the Borel \(\sigma\)-algebra of \(\RR\) that is generated by the usual open set topology.
\end{notation}

\begin{gap} This section needs to be continued with product spaces, the upper/lower continuity of measures, the construction of the Lebesgue measure, the Caratheodory theorem, and the Dynkin uniqueness theorem.\end{gap}

\subsection{Probability spaces}

\begin{definition}
Let \((\Omega, \col{f}, \p)\) be a measure space. If \(\p(\col{f}) = 1\), we call \(\p\) a \defname{probability measure}, we call \((\Omega, \col{f}, \p)\) a \defname{probability space}, and we call an element of \(\col{f}\) an \defname{event}.
\end{definition}

\begin{notation}
It is common to use a shorthand notation for events. We often write \(\{\text{condition}\}\) to mean \(\{\omega \in \Omega\;:\;\text{condition}\}\) and we often write \(\p(\text{condition})\) to mean \(\p\left(\{\omega \in \Omega\;:\;\text{condition}\}\right)\).
\end{notation}

\begin{theorem}\label{thm.nestedconvergence}
Let \(\{A_n\}_{n \in \NN}\) be a sequence of events.
\begin{enumerate}
\item
If \(A_1 \subset A_2 \subset A_3 \subset \cdots\) then \(\lim_{n \to \infty} \p(A_n) = \p\left(\cup_{n=1}^\infty A_n\right)\).
\item
If \(A_1 \supset A_2 \supset A_3 \supset \cdots\) then \(\lim_{n \to \infty} \p(A_n) = \p\left(\cap_{n=1}^\infty A_n\right)\).
\end{enumerate}
\end{theorem}

\begin{proof}\ 
\begin{enumerate}
\item
Let \(A = \cup_{n=1}^\infty A_n\). Let \(B_1 = A_1\) and for \(n > 1\) let \(B_n = A_n \setminus A_{n-1}\). Then the \(B_n\) are disjoint.
Now, \(A_n = \cup_{i=1}^n B_i\), so \(\p(A_n) = \sum_{i=1}^n \p(B_n)\). Also \(A = \cup_{n=1}^\infty B_n\) so \(\p(A) = \sum_{i=1}^\infty \p(B_i)\). This sum converges because the partial sums \(\sum_{i=1}^n \p(B_i)\) are monotonically increasing and bounded above by \(1\). The remainder \(R_n = \sum_{i = n+1}^\infty \p(B_n)\) has \(R_n \to 0\) as \(n \to \infty\).
Finally, we see that \(\p(A) - \p(A_n) = R_n \to 0\) as \(n \to \infty\). Thus \(\p(A_n) \to \p(A)\) as \(n \to \infty\).

\item
This follows from the previous part by taking compliments.
\end{enumerate}
\end{proof}

\begin{definition}
If \((A_i)_{i \in \NN}\) are a sequence of events, ``\(A_i\) occurring \defname{infinitely often}'' is the event
\[\{A_i \text{ i.o.}\} = \bigcap_{m \in \NN} \;\bigcup_{n \geq m} A_n\]
and ``\(A_i\) occurring \defname{eventually}'' is the event
\[\{A_i \text{ eventually}\} = \bigcup_{m \in \NN} \;\bigcap_{n \geq m} A_n.\]
\end{definition}

\begin{remark}
\(\{A_i \text{ i.o.}\}\) is the set of elements \(\omega \in \Omega\) that belong in infinitely many of the events \((A_i)\) and \(\{A_i \text{ eventually}\}\) is the set of elements \(\omega \in \Omega\) that belong in all \((A_i)\) after a certain point. We note that \(\{A_i \text{ eventually}\} \subset \{A_i \text{ i.o.}\}\).
\end{remark}

\begin{definition}
The \defname{characteristic function} of an event \(A\) in \(\Omega\) is the function \(\ind{A}:\Omega \to \{0, 1\}\) that has \(\ind{A}(t) = 1\) if \(t \in A\) and \(\ind{A}(t) = 0\) if \(t \not\in A\).
\end{definition}

\begin{theorem}\label{thm.eventsio.charf}
Let \((A_i)_{i \in \NN}\) be a sequence of events.
\begin{enumerate}
\item
\(\ind{\{A_i \text{ i.o.}\}} = \limsup_{i \to \infty} \ind{A_i}\).
\item
\(\ind{\{A_i \text{ eventually}\}} = \liminf_{i \to \infty} \ind{A_i}\).
\end{enumerate}
\end{theorem}

\begin{proof}\ 
\begin{enumerate}
\item
Suppose \(\ind{\{A_i \text{ i.o.}\}}(t) = 1\). Then \(t \in \{A_i \text{ i.o.}\}\), so \(t\) is in infinitely many \(A_i\); call these \(A_{\alpha_i}\). Then \(\ind{A_{\alpha_i}}(t) = 1\) for all \(i\), so \(\lim_{i \to \infty} \ind{A_{\alpha_i}}(t) = 1\) and therefore \(\limsup_{i \to \infty} \ind{A_{i}}(t) = 1\).

Now suppose instead \(\limsup_{i \to \infty} \ind{A_{i}}(t) = 1\). Then there exists a subsequence \(\{A_{\alpha_i}\}\) of \(\{A_i\}\) and an \(N \in \NN\) so that \(\left|\ind{A_{\alpha_i}}(t) - 1 \right| < \frac{1}{2}\) for all \(i \geq N\). Then for these \(i\) it must be that  \(\ind{A_{\alpha_i}}(t) = 1\) so \(t \in A_{\alpha_i}\). Thus \(t\) is in infinitely many \(A_i\) so \(t \in \{A_i \text{ i.o.}\}\) and \(\ind{\{A_i \text{ i.o.}\}}(t) = 1\).

\item
Suppose \(\ind{\{A_i \text{ eventually}\}}(t) = 1\). Then \(t \in \{A_i \text{ eventually}\}\), so there exists an \(N \in \NN\) such that \(t\) is in all \(A_i\) for \(i \geq N\). For these \(i\) we have \(\ind{A_i}(t)=1\), so \(\lim_{i \to \infty} \ind{A_i}(t) = 1\). In particular \(\liminf_{i \to \infty} \ind{A_i}(t) = 1\).

Now suppose instead \(\liminf_{i \to \infty} \ind{A_i}(t) = 1\). Then there is no subsequence \(\{A_{\alpha_i}\}\) of \(\{A_i\}\) with \(\lim_{i \to \infty}\{A_{\alpha_i}\}(t) =0\), so there are finitely many \(A_i\) that do not contain \(t\). It follows that there exists an \(N \in \NN\) such that for all \(i \geq N\), \(t \in A_i\). Therefore \(t \in \{A_i \text{ eventually}\}\) and \(\ind{\{A_i \text{ eventually}\}}(t) = 1\).
\end{enumerate}
\end{proof}

\begin{gap}
The above proof could probably be shortened.
\end{gap}

\begin{theorem}
Any events \(A_n\) have \(\p(A_n \text{ i.o.}) \geq \limsup_{n\to\infty} \p(A_n)\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{gap} Add the Borel-Cantelli Lemma and the definition of the Borel \(\sigma\)-algebra on \(\RR\) here.\end{gap}

\begin{definition}
Let \((\Omega, \col{f}, \p)\) be a probability space. Events \((A_i)_{i \in I} \in \col{f}\) are called \defname{independent} if for any finite \(J \subset I\) we have \(\p\left(\cap_{j \in J} A_j\right) = \prod_{j \in J} \p\left(A_j\right)\).
\end{definition}

\begin{theorem}
If \((A_n)\) are independent and for all \(n\) we define either \(B_n = A_n\) or \(B_n = A_n^C\) then all \((B_n)\) are independent.
\end{theorem}

\begin{gap}
Prove the above theorem.
\end{gap}

\begin{gap}
Add the Goldberg conjecture discussion here, as an example.
\end{gap}

\begin{definition}
Let \((\Omega, \col{f}, \p)\) be a probability space. Collections \((\col{A}_i) \subset \col{f}\) are called \defname{independent} if for any sequence \((A_i)\in\col{f}\) with \(A_i \in \col{A}_i\) the \((A_i)\) are independent.
\end{definition}

\begin{gap}
Add examples of independent and non-independent collections.
\end{gap}


\subsection{Random variables}

\begin{definition}
Let \((\Omega_1, \col{f}_1, \mu_1)\) and \((\Omega_2, \col{f}_2, \mu_2)\) be measure spaces. A function \(g: \Omega_1 \to \Omega_2\) is called \defname{measurable} if for all \(A \in \col{f}_2\) we have \(g^{-1}(A) \in \col{f}_1\).
\end{definition}

\begin{remark}
This is similar to the topological definition of continuous functions.
\end{remark}


\begin{definition}
A \defname{random variable} on a probability space \((\Omega, \col{f}, \p)\) is a measurable function \(\rv{X}:\Omega \to \RR\), where we use the usual Borel \(\sigma\)-algebra as the \(\sigma\)-algebra on \(\RR\).
\end{definition}

\begin{notation}
When writing down an event involving a random variable \(\rv{X}\), it is common to write \(\rv{X}\) when we really mean \(\rv{X}(\omega)\). For example, we might write \(\p(\rv{X} = 1)\) to mean \(\p\left(\left\{\omega \in \Omega : \rv{X}(\omega) = 1\right\}\right)\).
\end{notation}

\begin{gap}
The notation remark above could be made more clear.
\end{gap}

\begin{gap}
Add some examples of random variables.
\end{gap}

\begin{remark}
We can form an equivalence relation on random variables by writing \(\rv{X} \sim \rv{Y}\) if \(\p(X=Y)=1\), that is, if \(\{\omega \in \Omega : \rv{X}(\omega) \neq \rv{Y}(\omega)\}\) has measure zero. Equivalent random variables have the same probabilistic properties. Some may prefer to think of ``random variables'' as these equivalence classes of functions. In this sense, it doesn't make sense to ask what a random variable's value is for a certain \(\omega \in \Omega\) with \(\p(\{\omega\}) = 0\); it only makes sense to look at the random variable's values on events with measure greater than zero.
\end{remark}

\begin{definition}
The \defname{distribution function} of a random variable \(\rv{X}\) is the function \(F_\rv{X}: \RR \to \RR\) given by \(F(t) = \p(X \leq t)\).
\end{definition}

We immediately see that distribution functions are monotonically increasing and right continuous (i.e., \(\lim_{t\to s^+} F_\rv{X}(t) = F_\rv{X}(s)\) for all \(s \in [0,1)\)). Also, \(F_\rv{X}\) has a jump discontinuity at \(s\) if and only if \(\p(\rv{X} = s) > 0\) (and if so, the jump is of size \(\p(\rv{X} = s)\)). There is an interesting theorem that provides something like a converse.

\begin{theorem}
Let \(F:\RR \to [0,1]\) be monotonically increasing and right continuous (i.e., \(\lim_{t\to s^+} F(t) = F(s)\) for all \(s \in [0,1))\). Then there exists a probability space \((\Omega, \col{f}, \p)\) and a random variable \(\rv{X}:\Omega \to \RR\) such that \(F = F_\rv{X}\).
\end{theorem}

\begin{gap}
Add the two theorem proofs.
\end{gap}

\begin{definition}
Given random variables \((\rv{X}_i)\) let \(\col{A}_i = \{{\rv{X}_i}^{-1}(A) : A \in \bor\}\). We say that \((\rv{X}_i)\) are \defname{independent} if \((\col{A}_i)\) are independent. (Here, \(\bor\) is the usual Borel \(\sigma\)-algebra on \(\RR\).)
\end{definition}

\begin{theorem}
Let \((\rv{X}_i)\) be random variables and let \(\col{A}_i = \left\{\left\{\rv{X}_i \leq t\right\} : t \in \RR\right\}\). Suppose \((\col{A}_i)\) are independent. Then \((\rv{X}_i)\) are independent.
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{definition}
The \defname{joint distribution function} of random variables \(\rv{X}_1, \ldots, \rv{X}_n\) is the function \(F:\RR^n \to \RR\) given by \[F\left(\langle t_1, \ldots, t_n\rangle\right) = \p \left(\rv{X}_i \leq t_i \;\forall\; i \in \{0, \ldots, n\}\right).\]
\end{definition}

\begin{gap}
Add notes on stochastic domination, percolation, etc.
\end{gap}

\subsection{Sequences of random variables}

\begin{definition}
Let \(\rv{X}_n, \rv{X} : \Omega\to \RR\) be random variables. We say that \(\{\rv{X}_n\}\) \defname{converges almost surely} to \(\rv{X}\) and write \(\rv{X}_n \asto \rv{X}\) if \(\p\left(\left\{\omega \in \Omega : \rv{X}_n(\omega) \to \rv{X}(\omega)\right\}\right) = 1\). We say that \(\{\rv{X}_n\}\) \defname{converges in probability} to \(\rv{X}\) and write \(\rv{X}_n \pto \rv{X}\) if for all \(\epsilon>0\) we have \(\p\left(\left|\rv{X}_n-\rv{X}\right| < \epsilon\right) \to 1\) as \(n \to \infty\).
\end{definition}

\begin{definition}\label{def.disto}
Let \(\rv{X}_n : \Omega_n \to \RR\) and \(\rv{X}:\Omega \to \RR\) be random variables. We say that \(\{\rv{X}_n\}\) \defname{converges in distribution} to \(\rv{X}\) and write \(\rv{X}_n \disto \rv{X}\) if \(F_{\rv{X}_n}(t) \to F_\rv{X}(t)\) for all \(t\in\RR\) such that \(\p\left(\rv{X}=t\right) = 0\). 
\end{definition}

\begin{theorem}
Let \(\rv{X}_n, \rv{X} : \Omega\to \RR\) be random variables.
\begin{enumerate}
\item
If \(\rv{X}_n \asto \rv{X}\) then \(\rv{X}_n \pto \rv{X}\).
\item
If \(\rv{X}_n \pto \rv{X}\) then \(\rv{X}_n \disto \rv{X}\).
\end{enumerate}
\end{theorem}

\begin{proof}\ 
\begin{enumerate}
\item
Suppose \(\rv{X}_n \asto \rv{X}\). Let \(S = \left\{\omega \in \Omega : \rv{X}_n(\omega) \to \rv{X}(\omega)\right\}\). Then \(\p(S) = 1\). 

Let \(\epsilon > 0\) and let \(A_n^\epsilon = \{|\rv{X}_n-X| < \epsilon\}\). For all \(\omega \in S\), there exists an \(N\) such that \(|\rv{X}_n(\omega) - X(\omega)| < \epsilon\) for all \(n \geq N\); in other words, if \(n \geq N\) then \(\omega \in A_n^\epsilon\). Thus \(S \subset \{A_n^\epsilon \text{ eventually}\}\). Since \(\p(S) = 1\), we must have \(\p\left(\{A_n^\epsilon \text{ eventually}\}\right) = 1\).

...


%
%\(\rv{X}_n \asto \rv{X}\) if and only if \(\p\left(A_n^\epsilon \text{ i.o.}\right) = 1\) for all \(\epsilon > 0\).
%
%\(\rv{X}_n \pto \rv{X}\) if and only if \(\p\left(A_n^\epsilon\right) \to 1\) as \(n \to \infty\) for all \(\epsilon > 0\).


\item
...
\end{enumerate}
\end{proof}

\begin{gap}
Finish the proof. Add counterexamples to show that converses not true.
\end{gap}

\begin{theorem}{(Uniqueness)}
\begin{enumerate}
\item
Suppose \(\rv{X}_n \pto \rv{X}\) and \(\rv{X}_n \pto \rv{Y}\). Then \(\p(\rv{X}=\rv{Y}) = 1\).
\item
Suppose \(\rv{X}_n \disto \rv{X}\) and \(\rv{X}_n \disto \rv{Y}\). Then \(F_\rv{X} = F_\rv{Y}\).
\end{enumerate}
\end{theorem}

\begin{proof}\ 
\begin{enumerate}
\item
Let \(\epsilon > 0\).
We have \(|\rv{X}-\rv{Y}| \leq |\rv{X}_n-\rv{X}| + |\rv{X}_n - \rv{Y}|\), so if \(|\rv{X}-\rv{Y}|\geq \epsilon\) then either \(|\rv{X}_n - \rv{X}| \geq \frac{\epsilon}{2}\) or \(|\rv{X}_n - \rv{Y}| \geq \frac{\epsilon}{2}\). Thus \[\{|\rv{X}-\rv{Y}|\geq \epsilon\} \subset \left\{|\rv{X}_n-\rv{X}| \geq \frac{\epsilon}{2}\right\} \cup \left\{|\rv{X}_n - \rv{Y}| \geq \frac{\epsilon}{2}\right\}\]
so
\[\p(|\rv{X}-\rv{Y}|\geq \epsilon) \leq \p\left(|\rv{X}_n-\rv{X}| \geq \frac{\epsilon}{2}\right) + \p\left(|\rv{X}_n - \rv{Y}| \geq \frac{\epsilon}{2}\right) \to 0\]
as \(n \to \infty\). Therefore \(\p(|\rv{X}-\rv{Y}| < \epsilon) = 1\).

Take \(A_1 = \{|\rv{X}-\rv{Y}| < 1\}\) and for \(n \geq 2\) take \(A_n = A_{n-1} \cap \{|\rv{X}-\rv{Y}| < \frac{1}{n}\}\). Then all \(\p(A_n) = 1\) and \(A_1 \supset A_2 \supset A_3 \supset \cdots\). Let \(A = \cap_n A_n\). Then by Theorem \ref{thm.nestedconvergence}, \(\p(A) = 1\). Also for all \(\omega \in A\) we must have \(\rv{X}(\omega) = \rv{Y}(\omega)\). We conclude that \(\p(\rv{X}=\rv{Y}) = 1\).

\item
From Definition \ref{def.disto} it is immediately clear that \(F_\rv{X} = F_\rv{Y}\) except where \(F_\rv{X}\) is discontinuous. This can happen for at most countably many \(t \in \RR\) (because at each discontinuity there exists a unique rational not in the range of \(F_\rv{X}\)). Then \(\{s \in \RR:F_\rv{X}(s) = F_\rv{Y}(s)\}\) is dense everywhere in \(\RR\). Thus for any \(t\in\RR\) where \(F_\rv{X}\) is discontinuous, there exist \(t_n\in \RR\) with \(t_n \to t\) and \(t_n > t\) and \(F_\rv{X}(t_n) = F_\rv{Y}(t_n)\) for all \(n\). Recalling that \(F_\rv{X}\) and \(F_\rv{Y}\) are right-continuous, we must have \(F_\rv{X}(t) = \lim_{n \to \infty} F_\rv{X}(t_n) = \lim_{n \to \infty} F_\rv{Y}(t_n) = F_\rv{Y}(t) \). This shows that \(F_\rv{X} = F_\rv{Y}\) everywhere.
\end{enumerate}
\end{proof}

\begin{theorem}
Suppose \(\rv{X}_n \pto \rv{X}\). Then there exists a subsequence \(\{\rv{X}_{n_k}\}\) of \(\{\rv{X}_n\}\) such that \(\rv{X}_{n_k} \asto \rv{X}\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}


\begin{theorem}{(Skorohod's Representation Theorem)}\ 

Suppose \(\rv{X}_n \disto \rv{X}\). Then there exist \(\rv{Y}_n, \rv{Y}\) such that all \(F_{\rv{X}_n} = F_{\rv{Y}_n}\) and \(F_\rv{X} = F_\rv{Y}\) and \(\rv{Y}_n \asto \rv{Y}\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}


\subsection{Expectation values}

\begin{definition}
A \defname{simple random variable} is a random variable \(\rv{X}:\Omega \to \RR\) such that \(\rv{X} = \sum_{i=1}^N y_i \ind{A_i}\) for some \(y_i \in \RR\) and disjoint \(A_i \subset \Omega\) and \(N \in \NN\).
\end{definition}

\begin{theorem}
For any random variable \(\rv{X}\) there exist simple random variables \(\rv{X}_n\) such that for all \(n\) we have \(\rv{X}_n \leq \rv{X}_{n+1}\) everywhere and \(\rv{X}_n \to \rv{X}\).
\end{theorem}

\begin{definition}
Let \(\rv{X} = \sum_{i=1}^N y_i \ind{A_i}\) be a simple random variable (with disjoint \(A_i\)). The \defname{expectation} of \(\rv{X}\) is \(\ex \rv{X} = \sum_{i=1}^N y_i \p(A_i)\).
\end{definition}

\begin{theorem}\ 
\begin{enumerate}
\item If \(\rv{X} \geq 0\) everywhere then \(\ex \rv{X} \geq 0\).
\item For simple random variables \(\rv{X}\) and \(\rv{Y}\) and \(a \in \RR\) we have \(\ex a \rv{X} = a \ex \rv{X}\) and \(\ex (X+Y) = \ex X + \ex Y\) (linearity).
\item If \(\rv{X} \leq \rv{Y}\) everywhere then \(\ex\rv{X} \leq \ex\rv{Y}\) (monotonicity).
\end{enumerate}
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

Now that we have defined the expectation for simple random variables, we want to extend the definition to all random variables. First we extend it to positive random variables.

\begin{definition}
Let \(\rv{X}: \Omega \to \RR\) be a random variable with \(\rv{X} > 0\) everywhere. Define the \defname{expectation} of \(\rv{X}\) to be
\[\ex \rv{X} = \sup \left\{\ex \rv{Y} : \rv{Y} \text{ is a simple random variable with } \rv{Y} \leq \rv{X} \text{ everywhere} \right\}.\]
\end{definition}

\begin{theorem}{(Monotone Convergence Theorem)}\ 

Suppose \(\rv{X}_n \to \rv{X}\) and always \(\rv{X}_{n+1} \geq \rv{X}_{n}\) everywhere. Then \(\ex \rv{X}_n \to \ex \rv{X}\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{gap}
Discuss the expectation for random variables in general, not just non-negative random variables.
\end{gap}

\begin{notation}
Instead of \(\ex \rv{X}\) we sometimes write \(\int \rv{X} d\p\) or \(\int \rv{X}(\omega) d\p(\omega)\).
\end{notation}

\begin{definition}
We say that \(\rv{X}_n\) \defname{converge in expectation} to \(\rv{X}\) and write \(\rv{X}_n \exto \rv{X}\) if \(\ex |\rv{X}_n - \rv{X}| \to 0\) as \(n \to \infty\).
\end{definition}

\begin{theorem}{(Markov's Inequality)}
If \(\rv{X} > 0\) everywhere then \(\p(\rv{X} \geq t) \leq \frac{\ex \rv{X}}{t}\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{theorem}{(Fatou's Lemma)}
If \(\rv{X}_n \geq 0\) everywhere then
\begin{enumerate}
\item
\(\liminf \ex \rv{X}_n \geq \ex \liminf \rv{X}_n\)
\item
\(\limsup \ex \rv{X}_n \leq \ex \limsup \rv{X}_n\).
\end{enumerate}
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{theorem}{(Dominated Convergence Theorem)}
Let \(\rv{X}_n, \rv{X}, \rv{Y}\) be random variables with all \(|\rv{X}_n| \leq \rv{Y}\) and \(\rv{X}_n \asto \rv{X}\) and \(\ex \rv{Y}\) is finite. Then \(\ex \rv{X} = \lim_{n \to \infty} \ex \rv{X}_n\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{gap}
Add the example theorem in graph theory.
\end{gap}

\begin{theorem}
If \(\rv{X}_i\) are independent random variables then \(\ex \prod_i \rv{X}_i = \prod_i \ex \rv{X}_i\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{theorem}
If \(\rv{X}_i\) are random variables each with finite expectation then \(\ex \prod_i \rv{X}_i = \prod_i \ex \rv{X}_i\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{theorem}
Random variables \(\rv{X}_i\) are independent if and only if for all bounded functions \(f_i\), \(\ex \prod_i f_i(\rv{X}_i) = \prod_i \ex f_i(\rv{X}_i)\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{definition}
Random variables \(\rv{X}\) and \(\rv{Y}\) are called \defname{uncorrelated} if \(\ex (\rv{X} \rv{Y}) = (\ex \rv{X})(\ex \rv{Y})\).
\end{definition}

\begin{theorem}{(Law of Large Numbers)}\ 

Let \(\rv{X}_n\) be independent identically-distributed random variables each with \(\ex |\rv{X}_n|\) finite. Then \[\frac{1}{n} \sum_{i=1}^n \rv{X}_i \asto \ex \rv{X}_1\] as \(n \to \infty\).
\end{theorem}

\begin{gap}
Prove the theorem.
\end{gap}

\begin{gap}
Add sections on density, discrete random variables, the central limit theorem, Poisson and Gaussian random variables, characteristic functions, the inversion formula, the continuity theorem, tight random variables, covariance, jointly normal random variables, the multi-dimensional central limit theorem, and branching processes.
\end{gap}

\subsection{Conditional probabilities and expectations}

\begin{definition}
Let \(A,B\in\Omega\) be events with \(\p(B)\neq 0\). The \defname{probability of \(A\) conditional on \(B\)} is
\[\p(A|B) = \frac{\p(A \cap B)}{\p(B)}.\]
\end{definition}

\begin{remark}
\(A\) and \(B\) are independent if and only if \(\p(A|B) = \p(A)\).
\end{remark}

\begin{remark}
\(\p(\cdot|B)\) is a probability measure on \(\Omega\).
\end{remark}

\begin{definition}
Let \((\Omega, \col{f}, \p)\) be a probability space and let \(\rv{X}:\Omega \to \RR\) be a measurable function with respect to \(\col{f}\). Let \(\col{g} \subset \col{f}\) be a smaller \(\sigma\)-algebra in \(\col{f}\). Let \(\rv{U}:\Omega\to\RR\) be measurable with respect to \(\col{g}\). We say that \(\rv{U}\) is a \defname{version of \(\ex(\rv{X}|\col{g})\)} if \(\ex \rv{X} \ind{A} = \ex \rv{U} \ind{A}\) for all \(A \in \col{g}\).
\end{definition}

\begin{remark}
We can choose one such \(\rv{U}\) and call it \(\ex(\rv{X}|\col{g})\).
\end{remark}

\begin{remark}
If \(\rv{X}\) is measurable with respect to \(\col{g}\) then \(\rv{X}\) is a version of \(\ex(\rv{X}|\col{g})\).
\end{remark}

\begin{definition}
Let \(\rv{X}, \rv{Y} : \Omega \to \RR\) be random variables and let \(\sigma(\rv{Y})\) be the random variable generated by \(\rv{Y}\). Then we define \(\ex(\rv{X}|\rv{Y}) = \ex(\rv{X}|\sigma(\rv{Y}))\).
\end{definition}






















\end{document}
